---
title: "Homework 6"
output: 
  rmdformats::robobook:
    css: "homework.css"
    highlight: pygments
link-citations: yes
---

**Due:** Friday, October 20, 11:00am on Canvas

**Instructions:** 

* Download the [HW 6 template](https://sta279-f23.github.io/homework/hw_6_template.qmd), and open the template (a Quarto document) in RStudio. 
* Put your name in the file header
* Click `Render`
* Type all code and answers in the document (using `###` for section headings and `####` for question headings)
* Render early and often to catch any errors!
* When you are finished, submit the final rendered HTML to Canvas

**Code guidelines:**

* If a question requires code, and code is not provided, you will not receive full credit
* You will be graded on the quality of your code. In addition to being correct, your code should also be easy to read
  * No magic numbers
  * Use descriptive names for your variables
  * Set seeds where needed
  * Comment code
  * If a block of code is being called multiple times, put it in a function
  
**Resources:** In addition to the class notes and activities, I recommend reading the following resources:

* [Chapter 4](https://mdsr-book.github.io/mdsr3e/04-dataI.html) and [Chapter 6](https://mdsr-book.github.io/mdsr3e/06-dataII.html) in *Modern Data Science with R*
* [Chapter 4](https://r4ds.hadley.nz/data-transform) and [Chapter 6](https://r4ds.hadley.nz/data-tidy) in *R for Data Science* (2nd edition)



## Baseball data

The `Teams` data in the `Lahman` package contains information on professional baseball teams since 1871. 

:::{.question}
#### Question 1

Using the `Teams` data, create a plot of the number of home runs scored (`HR`) and allowed (`HRA`) by the Chicago Cubs (`teamID` CHN) in each season. Your plot should look like close to this:

```{r, echo=F, message=F, warning=F}
library(Lahman)
library(tidyverse)

Teams |>
  filter(teamID == "CHN") |>
  select(yearID, HR, HRA) |>
  pivot_longer(-yearID,
               names_to = "team",
               values_to = "runs") |>
  mutate(team = ifelse(team == "HR", "Cubs", "Opponent")) |>
  ggplot(aes(x = yearID, y = runs, color = team)) +
  geom_point() +
  geom_line() +
  labs(x = "Year", y = "Runs", color = "Team") +
  theme_bw()
```

You may use whichever R functions you like to create the plot, but the axes and legend should be labeled as in the plot above.

:::


## Practice pivoting

:::{.question}
#### Question 2

The code below creates a data frame called `df1`:

```{r}
df1 <- data.frame(
  grp = c("A", "A", "B", "B"),
  sex = c("F", "M", "F", "M"),
  meanL = c(0.225, 0.47, 0.325, 0.547),
  sdL = c(0.106, 0.325, 0.106, 0.308),
  meanR = c(0.34, 0.57, 0.4, 0.647),
  sdR = c(0.0849, 0.325, 0.0707, 0.274)
)

df1
```

Using `pivot_longer` and `pivot_wider`, reshape `df1` to produce the following output:

```{r, echo=F}
df1 |>
  pivot_longer(-c(grp, sex),
               names_to = "stat", values_to = "value") |>
  pivot_wider(id_cols = grp,
              names_from = c(sex, stat),
              names_sep = ".",
              values_from = value)
```

Remember that the `?` provides documentation in R. For example, running `?pivot_wider` in the console gives helpful information about the `pivot_wider` function.
:::

## Practice pivoting

The code below creates a data frame called `df2`:

```{r}
df2 <- data.frame(id = rep(c(1, 2, 3), 2),
                  group = rep(c("T", "C"), each=3),
                  vals = c(4, 6, 8, 5, 6, 10))

df2
```

An analyst wants to calculate the pairwise differences between the Treatment (T) and Control (C) values for each individual in this dataset. They use the following code:

```{r}
Treat <- filter(df2, group == "T")
Control <- filter(df2, group == "C")
all <- mutate(Treat, diff = Treat$vals - Control$vals)
all
```


:::{.question}
#### Question 3

Verify that this code works for this example and generates the correct values of -1, 0, and -2. Describe two problems that might arise if the data set is not sorted in a particular order or if one of the observations is missing for one of the subjects.
:::

:::{.question}
#### Question 4

Provide an alternative approach to generate the `diff` variable, using `group_by` and `summarize` to produce the following output:

```{r, echo=F}
df2 |>
  group_by(id) |>
  summarize(diff = vals[group == "T"] - vals[group == "C"])
```

:::

:::{.question}
#### Question 5

Provide an alternative approach to generate the `diff` variable that uses `pivot_wider` and `mutate` to produce the following output:

```{r, echo=F}
df2 |>
  pivot_wider(id_cols = id,
              names_from = group,
              values_from = vals) |>
  mutate(diff = T - C)
```

:::


## Analyzing simulation results

Your friend ran simulations to compare three different methods for calculating 95% confidence intervals for $\beta_1$ in a linear regression model. Their data is store in the file `sim_results.csv`, which can be imported into R as follows:

```{r}
sim_results <- read.csv("https://sta279-f23.github.io/homework/sim_results.csv")
```


The `sim_results` dataset is organized as follows:

* The first column, `n`, records the sample size
* For each method and sample size, your friend simulated 100 datasets and calculated a confidence interval. The results for method 1 are stored in columns `method1_1`, `method1_2`, ..., `method1_100` (with similar columns for methods 2 and 3). A 1 in a column indicates that the confidence interval captured $\beta_1$, while a 2 indicates the confidence interval did not capture $\beta_1$.

:::{.question}
#### Question 6

Plot the observed coverage for each method (the fraction of times the confidence interval contained $\beta_1$) as a function of sample size. Your final plot should look as close as possible to this:

```{r, echo=F}
sim_results |>
  mutate(across(-n, function(x) {2 - x})) |>
  rowwise() |>
  mutate(method1 = mean(c_across(contains("method1"))),
         method2 = mean(c_across(contains("method2"))),
         method3 = mean(c_across(contains("method3")))) |>
  select(n, method1, method2, method3) |>
  pivot_longer(-n, 
               names_to = "method",
               values_to = "coverage") |>
  ggplot(aes(x = n, y = coverage, color = method)) +
  geom_point() +
  geom_line() +
  labs(x = "Sample size (n)",
       y = "Coverage",
       color = "Method") +
  theme_classic()
```

You will likely need to use some combination of the following functions: `pivot_longer`, `mutate`, `select`, `across`, `c_across`, `rowwise`

:::


## Fitting linear regression models

The following code generates data $X_i \sim N(0, 1)$, $\varepsilon_i \sim N(0, 1)$, and 

$$Y_i = \beta_0 + \beta_1 X_i + \varepsilon_i$$

with $\beta_0 = \beta_1 = 1$.

```{r}
n <- 100
x <- rnorm(n)
noise <- rnorm(n)
y <- 1 + x + noise
```

To estimate $\beta_0$ and $\beta_1$, we can use the `lm` function:

```{r}
lm(y ~ x)
```

How are these estimated coefficients calculated? Mathematically, the vector of estimated coefficients $\widehat{\beta}$ is given by

$$\widehat{\beta} = (X^T X)^{-1} X^T Y$$
where $X$ is the *design matrix* and $Y$ is the vector of observed responses. For the model above,

$$\widehat{\beta} = \begin{bmatrix}
\widehat{\beta}_0 \\
\widehat{\beta}_1
\end{bmatrix} \hspace{0.5cm} X = \begin{bmatrix}
1 & x_1 \\
1 & x_2 \\
1 & x_3 \\
\vdots & \vdots \\
1 & x_n
\end{bmatrix} \hspace{0.5cm} Y = \begin{bmatrix}
y_1 \\
y_2 \\
y_3 \\
\vdots \\
y_n
\end{bmatrix}$$

How would we do this math in R? Since $X$ is a matrix, and $\widehat{\beta}$ and $Y$ are both vectors, we want to work with *matrices* and *vectors* in R!

* The vector `y` generated above already represents $Y$ (by default, vectors in R are column vectors)
* We need to create the design matrix $X$, by adding a column of 1s to the `x` simulated above. The following code creates $X$:

```{r, eval=F}
X_mat <- cbind(1, x)
```

* Matrix multiplication in R is done with `%*%`. For example, if `A` and `B` are conformal matrices/vectors, then $AB$ would be calculated with `A %*% B`
* Transposes are calculated with `t()`. E.g., $A^T$ is given by `t(A)`
* Matrix inverses are calculated with `solve()`. E.g., $A^{-1}$ is given by `solve(A)`


:::{.question}
#### Question 7

Using this information, write a function called `my_lm` which will calculate the estimated coefficients $\widehat{\beta}$ given a vector of responses `y` and a design matrix `X_mat`.
:::




