---
title: "Homework 8 solutions"
format: 
  html:
    embed-resources: true
editor: source
author: "Ciaran Evans"
---

Total: 21 pts

1. (2 pts)

```{r, message=F, warning=F}
library(rvest)
library(tidyverse)

read_html("https://en.wikipedia.org/wiki/The_Great_British_Bake_Off_(series_2)") |>
  html_elements(".mw-headline[id^='Episode_']") |>
  html_text2() 
```

**Grading:** 2 pts for correct output

2. (4 pts)

```{r}
episodes <- read_html("https://en.wikipedia.org/wiki/The_Great_British_Bake_Off_(series_2)") |>
  html_elements(".mw-headline[id^='Episode_']") |>
  html_text2() 

data.frame(episode = episodes) |>
  separate_wider_regex(episode, patterns = c(".+ ", episode = "\\d+", ": ", 
                                          episode_name = ".+", "\\("),
                       too_few = "align_start")
```

**Grading:** 2 pts for separate columns for episode name and episode number. 2 pts for removing the words in parentheses

3. (2 pts)

```{r}
episodes <- read_html("https://en.wikipedia.org/wiki/The_Great_British_Bake_Off_(series_4)") |>
  html_elements(".mw-headline[id^='Episode_']") |>
  html_text2() 

data.frame(episode = episodes) |>
  separate_wider_regex(episode, patterns = c(".+ ", episode = "\\d+", ": ", 
                                          episode_name = ".+", "\\("),
                       too_few = "align_start")
```

There are several extra episodes which shouldn't be there, and which don't appear to have names. This is because of several "masterclass" episodes on that wikipedia page, which weren't part of the main show.

**Grading:** 2 pts for a reasonable answer. If they didn't have an error, because of how they originally wrote the code, that is ok.

4. (4 pts)

```{r}
episodes <- read_html("https://en.wikipedia.org/wiki/The_Great_British_Bake_Off_(series_4)") |>
  html_elements(".mw-headline[id^='Episode_']") |>
  html_text2() 

data.frame(episode = episodes) |>
  filter(str_detect(episode, ":")) |>
  separate_wider_regex(episode, patterns = c(".+ ", episode = "\\d+", ": ", 
                                          episode_name = ".+", "\\s+\\("),
                       too_few = "align_start")
```

**Grading:** If they didn't have an error (because they wrote their code differently) that is ok. 1 pt for separate columns for episode name and episode number. 2 pts for removing the words in parentheses. 1 pt for 10 rows in the output.

5. (5 pts)

```{r, message=F, warning=F}
library(data.table)

episode_list <- list()
urls <- paste0("https://en.wikipedia.org/wiki/The_Great_British_Bake_Off_(series_", 
               2:13, ")")

for(i in 2:13){
  episodes <- read_html(urls[i-1]) |>
    html_elements(".mw-headline[id^='Episode_']") |>
    html_text2() 

  episode_list[[i-1]] <- data.frame(episode = episodes) |> 
    filter(str_detect(episode, ":")) |>
    separate_wider_regex(episode, patterns = c(".+ ", episode = "\\d+", ": ", 
                                          episode_name = ".+", "\\s+\\("),
                       too_few = "align_start") |>
    mutate(series = i)
}

episodes <- rbindlist(episode_list)
head(episodes)
```

**Grading:** 2 pts for iterating over series, 2 pts for separate episode number and name columns, 1 pt for series column

6. (2 pts)

```{r}
episodes |>
  count(episode_name) |>
  filter(n == 1) |>
  nrow()
```

**Grading:** 2 pts for correct answer. They can get 1 pt if they arrived at an incorrect answer but did something reasonable

7. (2 pts)

```{r}
episodes |>
  count(episode_name, sort=T)
```

**Grading:** Answers may vary, depending on whether they count "Cake" and "Cakes" as the same, whether "Biscuits" is the same as "Biscuits and traybakes", etc.

