---
title: "Lecture 1: Intro to simulation"
format: 
  revealjs:
    theme: theme.scss
editor: visual
editor_options: 
  chunk_output_type: console
---

## Warm-up question

**Problem:** 10 people are at a party, and all of them are wearing hats. They each place their hat in a pile; when they leave, they choose a hat at random. What is the probability at least one person selected the correct hat?

**Question:** Work with your neighbor to discuss the following question:

-   Without calculating probabilities, how could you design an experiment to estimate this probability?

## Designing an experiment

## Step 1: representing the hats

```{r, echo=T}
hats <- 1:10

hats
hats[3]
```

-   `hats` is a **vector**, containing the numbers 1 to 10
-   entries in a vector are accessed by their index

## Step 2: everyone draws a random hat

```{r, echo=T}
hats <- 1:10
randomized_hats <- sample(hats, size = 10, replace = FALSE)

hats
randomized_hats
```

-   The `sample` function creates a random sample from a vector
-   How many people selected their original hat?

## Step 3: check who got their original hat

```{r, echo=T, eval=F}
hats <- 1:10
randomized_hats <- sample(hats, size = 10, replace = FALSE)
```

```{r, echo=T}
hats
randomized_hats
```

```{r, echo=T}
hats == randomized_hats

# TRUE is 1, FALSE is 0
sum(hats == randomized_hats)

# did at least one person get their hat?
sum(hats == randomized_hats) > 0
```

## Code so far

```{r, echo=T, eval=F}
hats <- 1:10
randomized_hats <- sample(hats, size = 10, replace = FALSE)
```

```{r, echo=T}
sum(hats == randomized_hats) > 0
```

-   Is this a good estimate of the probability?

## Step 4: iteration

A `for` loop repeats code many times:

```{r, echo=T, eval=F}
nsim <- 10000 # number of simulations 
for(i in 1:nsim){
  
  
}
```

## Step 4: iteration

A `for` loop repeats code many times:

```{r, echo=T}
nsim <- 10000 # number of simulations
hats <- 1:10
results <- rep(NA, nsim) # vector to store results

for(i in 1:nsim){
  randomized_hats <- sample(hats, size = 10, replace = FALSE)
  results[i] <- sum(hats == randomized_hats) > 0
}

head(results)
```

## Step 4: iteration

A `for` loop repeats code many times:

```{r, echo=T}
nsim <- 10000 # number of simulations
hats <- 1:10
results <- rep(NA, nsim) # vector to store results

for(i in 1:nsim){
  randomized_hats <- sample(hats, size = 10, replace = FALSE)
  results[i] <- sum(hats == randomized_hats) > 0
}

mean(results)
```

-   What if I wanted to repeat the simulation, with a different number of people?

## Removing magic numbers

Without magic numbers:

```{r, echo=T}
nsim <- 10000 # number of simulations
M <- 10 # number of people
hats <- 1:M
results <- rep(NA, nsim) # vector to store results

for(i in 1:nsim){
  randomized_hats <- sample(hats, 
                            size = M, 
                            replace = FALSE)
  results[i] <- sum(hats == 
                      randomized_hats) > 0
}

mean(results)
```

-   Why did I get different results?

## Final code

```{r, echo=T, eval=F}
set.seed(3) # set a seed for reproducibility

M <- 10 # number of people at the party
hats <- 1:M # numbered hats
nsim <- 10000 # number of simulations
results <- rep(NA, nsim) # vector to store the results

for(i in 1:nsim){
  # hats are randomly assigned to each person
  randomized_hats <- sample(hats, M, replace = F)
  
  # did at least one person get their hat back?
  results[i] <- sum(randomized_hats == hats) > 0
}

mean(results)
```

## Summary of coding practices

-   avoid magic numbers
-   set a seed for reproducibility
-   use meaningful names
-   add comments

## Class activity

Work with a neighbor on the class activity (link below and on the course website):

<https://sta279-s24.github.io/class_activities/ca_lecture_1.html>
