---
title: "Lecture 12: Data wrangling"
format: 
  revealjs:
    theme: theme.scss
editor: source
execute: 
  echo: true
editor_options: 
  chunk_output_type: console
---

## Last time

* `filter`: choose certain rows
* `summarize`: calculate summary statistics
* `group_by`: group rows together
* `mutate`: create new columns

## Data for today

* Data on professional baseball teams between 1871 and 2022
* 3015 rows and 48 columns
* Each row represents one year (season) for one team
* Variables include:
  * `yearID`: Year
  * `franchID`: Franchise
  * `W`: Wins
  * `L`: Losses
  
## Data for today

* Variables include:
  * `yearID`: Year
  * `franchID`: Franchise
  * `W`: Wins
  * `L`: Losses
  
We want to know: which NY Mets general manager performed best between 1998 - 2018

## Making a plan

We want to know: which NY Mets general manager performed best between 1998 - 2018

**Question:** What steps could we take to answer this question?

```{r, include=F}
library(Lahman)
library(tidyverse)
```

## Step 0: Make the columns more manageable

There are 48 columns in the initial data! Let's only focus on the ones we care about:

```{r}
Teams |>
  select(yearID, franchID, W, L)
```

## Step 1: Focus on the Mets


```{r, eval=F}
Teams |>
  select(yearID, franchID, W, L) |>
  ...(franchID == "NYM")
```

What function do I use to choose only the rows corresponding to the Mets?


## Step 1: Focus on the Mets

```{r}
Teams |>
  select(yearID, franchID, W, L) |>
  filter(franchID == "NYM")
```

## Step 2: Focus on the Mets between 1998 and 2018

```{r, eval=F}
Teams |>
  select(yearID, franchID, W, L) |>
  filter(franchID == "NYM",
         ...)
```

How do I specify the range of years I want?

## Step 2: Focus on the Mets between 1998 and 2018

```{r}
Teams |>
  select(yearID, franchID, W, L) |>
  filter(franchID == "NYM",
         yearID >= 1998, yearID <= 2018)
```

## Step 3: Who was the GM?

* 1998 - 2003: Steve Phillips
* 2004: Jim Duquette
* 2005 - 2010: Omar Minaya
* 2011 - 2018: Sandy Alderson

How should we add this information to the data?

## Step 3: Who was the GM?

```{r}
Teams |>
  select(yearID, franchID, W, L) |>
  filter(franchID == "NYM",
         yearID >= 1998, yearID <= 2018) |>
  mutate(gm = case_when(
    yearID <= 2003 ~ "Phillips",
    yearID == 2004 ~ "Duquette",
    yearID <= 2010 ~ "Minaya",
    yearID <= 2018 ~ "Alderson"
  ))
```

## Step 4: Summarize performance

How do I calculate performance for *each* GM?

```{r}
Teams |>
  select(yearID, franchID, W, L) |>
  filter(franchID == "NYM",
         yearID >= 1998, yearID <= 2018) |>
  mutate(gm = case_when(
    yearID <= 2003 ~ "Phillips",
    yearID == 2004 ~ "Duquette",
    yearID <= 2010 ~ "Minaya",
    yearID <= 2018 ~ "Alderson"
  )) |>
  summarize(wpct = sum(W)/sum(W + L))
```

## Step 4: Summarize performance

```{r}
Teams |>
  select(yearID, franchID, W, L) |>
  filter(franchID == "NYM",
         yearID >= 1998, yearID <= 2018) |>
  mutate(gm = case_when(
    yearID <= 2003 ~ "Phillips",
    yearID == 2004 ~ "Duquette",
    yearID <= 2010 ~ "Minaya",
    yearID <= 2018 ~ "Alderson"
  )) |>
  group_by(gm) |>
  summarize(wpct = sum(W)/sum(W + L))
```

## Finally: arrange results

```{r}
Teams |>
  select(yearID, franchID, W, L) |>
  filter(franchID == "NYM",
         yearID >= 1998, yearID <= 2018) |>
  mutate(gm = case_when(
    yearID <= 2003 ~ "Phillips",
    yearID == 2004 ~ "Duquette",
    yearID <= 2010 ~ "Minaya",
    yearID <= 2018 ~ "Alderson"
  )) |>
  group_by(gm) |>
  summarize(wpct = sum(W)/sum(W + L)) |>
  arrange(desc(wpct))
```

## Class activity

[https://sta279-s24.github.io/class_activities/ca_lecture_12.html](https://sta279-s24.github.io/class_activities/ca_lecture_12.html)

