---
title: "Class activity"
output: 
  tufte::tufte_html:
    css: "lab.css"
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

## Airlines data

The `airlines` SQL database contains information on 48 million flights from the Bureau of Transportation Statistics (you have worked with a small subset of this data in the `nycflights13` package). Information from the database can be obtained through SQL queries. For example, the flights contains the following tables:

```{r, include=F}
library(tidyverse)
library(mdsr)
library(DBI)
db <- dbConnect_scidb("airlines")
flights <- tbl(db, "flights")
carriers <- tbl(db, "carriers")
planes <- tbl(db, "planes")
```

```{sql connection=db}
SHOW TABLES;
```

## Connecting to the database

Make sure the `mdsr` and `DBI` packages are installed, then run the following code in R:

```{r}
library(tidyverse)
library(mdsr)
library(DBI)
db <- dbConnect_scidb("airlines")
```

Let's try a quick query to see that it works:

```{r}
query <- "
SELECT name FROM carriers LIMIT 0,5;
"
dbGetQuery(db, query)
```

Here, we are querying SQL through R. We store the query in a string, and then use the `dbGetQuery` function to send that query to the database and return the results.


1. Fill in the query below to select the `tailnum` and `year` columns from the `planes` table. For safety, we are only going to return the first 5 rows.

```{r, eval=F}
query <- "
SELECT ... FROM ... LIMIT 0,5;
"
dbGetQuery(db, query)
```

```{r, eval=F, echo=F}
query <- "
SELECT tailnum, year FROM planes LIMIT 0,5;
"
dbGetQuery(db, query)
```

2. Using the `MIN()` function, find the oldest date of manufacture (`year`) in the `planes` table.

```{r, eval=F}
query <- "
SELECT ... FROM ... LIMIT 0,5;
"
dbGetQuery(db, query)
```

```{r, eval=F, echo=F}
query <- "
SELECT tailnum, MIN(year) FROM planes LIMIT 0,5;
"
dbGetQuery(db, query)
```

3. Using `WHERE`, find the `tailnum` of the oldest plane in the `planes` table.

```{r, eval=F}
query <- "
SELECT ... 
FROM ... 
WHERE ...
LIMIT 0,5;
"
dbGetQuery(db, query)
```

```{r, eval=F, echo=F}
query <- "
SELECT tailnum, year FROM planes WHERE year = 1956 LIMIT 0,5;
"
dbGetQuery(db, query)
```

Questions 2 and 3 required two separate queries: first to find the oldest plane, then to find the tail number. We can be more efficient by using a *subquery* to choose only the years equal to the smallest year

```{r}
query <- "
SELECT tailnum, year 
FROM planes 
WHERE year = (SELECT MIN(year) FROM planes)
LIMIT 0,5;
"
dbGetQuery(db, query)
```

4. Modify the above code to find the *newest* manufacturing date, and calculate the number of planes which share the newest manufacturing date. You can use `SUM(1)` or `COUNT(...)` to count rows.

```{r, echo=F, eval=F}
query <- "
SELECT year, SUM(1) AS N 
FROM planes 
WHERE year = (SELECT MAX(year) FROM planes)
LIMIT 0,10;
"
dbGetQuery(db, query)
```

```{r, echo=F, eval=F}
query <- "
SELECT year, COUNT(year) AS N 
FROM planes 
WHERE year = (SELECT MAX(year) FROM planes)
LIMIT 0,10;
"
dbGetQuery(db, query)
```

