---
---
---
title: "Class activity, January 17"
format: html
editor: visual
author: "Your name"
---

### The birthday problem

Suppose we have a class of 30 students. What is the probability that there is at least one shared birthday?

* Assume there are 365 days in a year
* Assume that each day is equally likely as a birthday
* Assume there are no multiple-birth siblings (e.g. twins, triplets, etc.) in the class

1. Create a vector, `days`, to store the days of the year (numbered 1 to 365)

```{r}

```


2. Use the `sample` function to randomly choose birthdays for the class (shared birthdays are possible, so use `replace=TRUE` in the function!)

```{r}

```


3. Check the number of distinct birthdays (hints: `unique` returns the unique elements of a vector; `length` returns the length of a vector). Are there 30 distinct birthdays?

```{r}

```


4. Use a `for` loop to repeat the experiment $n_{sim} = 10000$ times, making sure to store the results. What is the probability of at least one shared birthday?

```{r}

```


5. Repeat the simulation with different class sizes. How many students do we need for the probability of a shared birthday to be approximately 50%?

```{r}

```


