---
title: "Class Activity"
output: 
  tufte::tufte_html:
    css: "lab.css"
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

**Instructions:** Work with a neighbor to answer the following questions. To get started, download the [class activity template](https://sta279-s24.github.io/class_activities/ca_lecture_14_template.qmd) file.


# GDP data

Run the following code to import data on GDP for each country (you may need to install the `googlesheets4` package):

```{r, message=F, warning=F}
library(tidyverse)
library(googlesheets4)
gs4_deauth()
gdp_data <- read_sheet("https://docs.google.com/spreadsheets/d/1RctTQmKB0hzbm1E8rGcufYdMshRdhmYdeL29nXqmvsc/pub?gid=0")
```

Looking at the data, we can see that there is a column for the GDP in each year, and the first column contains country information:

```{r}
head(gdp_data)
```


1. Use the `pivot_longer` function to reshape the GDP data so it looks like this:

```{r, echo=F}
gdp_data |>
  rename(country = starts_with("Income")) |>
  pivot_longer(cols = -country,
               names_to = "year",
               values_to = "gdp",
               values_drop_na = T)
```

* There should be no NAs
* The first column should be renamed `country`
* There should be separate columns for `year` and `gdp`

# WHO data

The `tidyr` package contains a dataset called `who2` which contains information on tuberculosis counts for each country in each year, broken down by method of diagnosis, sex, and age group:

```{r}
who2
```

Columns like `sp_m_1524` contain the number of tuberculosis cases. The column names contain information about the method of diagnosis (`rel`, `sn`, `sp`, or `ep`), sex (`f` or `m`), and age group (`014` = 0-14 years, `1524` = 15-24 years, etc.).

Ideally, we would like these variables to each have their own column. 

2. Reshape the `who2` data so it looks like this:

```{r, echo=F}
who2 |>
  pivot_longer(cols = -c(country, year),
               names_to = c("diagnosis", "sex", "age_group"),
               names_sep = "_",
               values_to = "count",
               values_drop_na = T)
```

* There should be no NAs
* There should be separate columns for `diagnosis`, `sex`, and `age_group`
* The `count` column contains the number of cases

