---
title: "Class Activity"
output: 
  tufte::tufte_html:
    css: "lab.css"
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

**Instructions:** Work with a neighbor to answer the following questions, then we will discuss the activity as a class. To get started, download the [class activity template](https://sta279-s24.github.io/class_activities/ca_lecture_5_template.qmd) file.

## Writing a function to calculate the mean

The *mean* of numbers $x_1,...,x_n$ is their average: 

$$\frac{x_1 + x_2 + \cdots + x_n}{n}$$
In R, we can calculate the mean with the `mean` function. For example:

```{r}
mean(c(1,2,3))
```

1. Write your own mean function, called `my_mean`, that calculates the mean of a vector. The input to your function should be a vector, and the output should be the mean of the values in that vector. You may *not* use the built-in `mean` function in R, but you *may* use the `sum` and `length` functions.

2. Check that your function works by comparing the output with R's `mean` function for several different input vectors.

## Weighted means

A *weighted* mean is similar to the usual average, but now we add a *weight* to each value. The observations with greater weights contribute more to the weighted mean.

The weighted mean of $x_1,...,x_n$ with weights $w_1,...,w_n$ is 

$$\frac{w_1 x_1 + w_2 x_2 + \cdots + w_n x_n}{w_1 + w_2 + \cdots + w_n}$$
The usual arithmetic mean is a special case of the weighted mean with $w_1 = w_2 = \cdots = w_n = 1$.

3. Modify your `my_mean` function from Exercise 1 so that it now takes **two** inputs -- a vector of values `x`, and a vector of weights `w` -- and returns the *weighted* mean of `x` with weights `w`. *Hint*: If `x` and `w` are two vectors of the same length, then `x*w` is a vector created by multiplying each entry of `x` with the corresponding entry of `w`.

4. When we calculate a mean, we usually don't need weights. Modify your `my_mean` function from Exercise 3 so that the *default* weights are all 1.

5. Check that your function from Exercise 4 works by running the following:

```{r, eval=F}
# should be 2
my_mean(x = c(1, 2, 3))

# should be 2
my_mean(x = c(1, 2, 3), w = c(1, 1, 1))

# should be 1.5
my_mean(x = c(1, 2, 3), w = c(1, 1, 0))
```




