---
title: "Class activity"
output: 
  tufte::tufte_html:
    css: "lab.css"
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

## Weighted sums

In a previous assignment, you have computed weighted sums and means. If $x_1,...,x_n$ is a set of observations and $w_1,...,w_n$ a set of weights, then $\sum \limits_{i=1}^n w_i x_i$ is the **weighted** sum.

In R, `sum(w * x)` calculates the weighted sum, for vectors `w` (containing $w_1,...,w_n$) and `x` (containing $x_1,...,x_n$).

1. In R, the `crossprod` function, when applied to vectors, computes the [inner product](https://en.wikipedia.org/wiki/Dot_product) (aka dot product) of the two vectors. Use the `crossprod` function to compute a weighted sum, and compare the speed with `sum(w * x)`.

2. In question 1, does the relative speed change as we change the length of `x`?

## Concatenating strings

The code below generates 10 random strings, of length 50 characters each, and combines them together:

```{r, eval=F}
random_string <- function() {
  paste(sample(letters, 50, replace = TRUE), collapse = "")
}
strings <- replicate(10, random_string())

collapse <- function(xs) {
  out <- ""
  for (x in xs) {
    out <- paste0(out, x)
  }
  out
}

collapse(strings)
```

3. Write faster code to combine the strings, using the `paste` function and avoiding a for loop. Compare the speed of your code to the `collapse` function above.

## C++ and Rcpp

When we can't optimize our code any further in R, we sometimes turn to C++ (a different programming language, which is often faster). In a subsequent lecture, we will discuss how to re-write R code in C++ for performance increases. For now, we will get Rcpp (the package used to integrate C++ code into R) set up.

The installation instructions can be found [here](https://adv-r.hadley.nz/rcpp.html#introduction-24) and [here](https://teuder.github.io/rcpp4everyone_en/020_install.html#install-rcpp).

Follow the installation instructions, then run the following code in R to make sure everything was installed correctly:

```{r, eval=F}
Rcpp::cppFunction('int add(int x, int y, int z) {
  int sum = x + y + z;
  return sum;
}')

add(1, 2, 3)
```





